---
title             : "Changes in Locus of Control Orientation Across College Major Pathways
"
shorttitle        : "Locus of Control and Major Pathways"

author: 
  - name          : "Varsha Venkatesh"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "75 E River Pkwy, Minneapolis, MN 55455"
    email         : "venka288@umn.edu"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Moin Syed"
    affiliation   : "1"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "University of Minnesota"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["references.bib", "r-references.bib"]

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library("DescTools")
library("dplyr")
library("summarytools")
r_refs("r-references.bib")

dat <- read.csv("/Users/varshaschool/Documents/loc_pathways/Data/loc_pathways_simulated_data_2025-11-08.csv")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

Participants were `r nrow(dat)` undergraduate students enrolled at a large, public university in the United States. They ranged in age from `r min(dat$age_sim)` to `r max(dat$age_sim)` years old (*M* = `r mean(dat$age_sim)`, *SD* = `r sd(dat$age_sim)`). The majority of participants identified as `r Mode(dat$race_sim)` (`r (attr(Mode(dat$race_sim), "freq"))/(length(dat$race_sim)) * 100`%) and `r tolower(Mode(dat$gender_sim))` (`r (attr(Mode(dat$gender_sim), "freq"))/(length(dat$gender_sim)) * 100`%). Additionally, the majority (`r (attr(Mode(dat$us_born_sim), "freq"))/(length(dat$us_born_sim)) * 100`%) of participants were born in the United States. A detailed summary of participant demographics is available in Table 1.

```{r descriptive table, echo=FALSE}
#Gender
gender_n <- ""
gender_p <- ""
male_n <- sum(dat$gender_sim == "Male")
male_p <- round((male_n/nrow(dat) * 100), digits = 2)
female_n <- sum(dat$gender_sim == "Female")
female_p <- round((female_n/nrow(dat) * 100), digits = 2)
nb_n <- sum(dat$gender_sim == "Nonbinary")
nb_p <- round((nb_n/nrow(dat) * 100), digits = 2)

#Race
race_n <- ""
race_p <- ""
asian_n <- sum(dat$race_sim == "Asian/Pacific Islander")
asian_p <- round((asian_n/nrow(dat) * 100), digits = 2)
black_n <- sum(dat$race_sim == "Black/African American")
black_p <- round((black_n/nrow(dat) * 100), digits = 2)
latin_n <- sum(dat$race_sim == "Hispanic/Latine")
latin_p <- round((latin_n/nrow(dat) * 100), digits = 2)
multi_n <- sum(dat$race_sim == "Multiracial")
multi_p <- round((multi_n/nrow(dat) * 100), digits = 2)
native_n <- sum(dat$race_sim == "Native American")
native_p <- round((native_n/nrow(dat) * 100), digits = 2)
other_race_n <- sum(dat$race_sim == "Other")
other_race_p <- round((other_race_n/nrow(dat) * 100), digits = 2)

#US-Born
row_us_born_n <- ""
row_us_born_p <- ""
us_born_n <- sum(dat$us_born_sim == 1)
us_born_p <- round((us_born_n/nrow(dat) * 100), digits = 2)
not_us_born_n <- sum(dat$us_born_sim == 0)
not_us_born_p <- round((not_us_born_n/nrow(dat) * 100), digits = 2)

dem_table <- tribble(
    ~Characteristic, ~n, ~`%`,
    "Gender", gender_n, gender_p,
    "     Male", as.factor(male_n), as.factor(male_p),
    "     Female", as.factor(female_n), as.factor(female_p),
    "     Nonbinary", as.factor(nb_n), as.factor(nb_p),
    "Race", race_n, race_p,
    "     Asian/Pacific Islander", as.factor(asian_n), as.factor(asian_p),
    "     Black/African American", as.factor(black_n), as.factor(black_p),
    "     Hispanic/Latine", as.factor(latin_n), as.factor(latin_p),
    "     Multiracial", as.factor(multi_n), as.factor(multi_p),
    "     Native American", as.factor(native_n), as.factor(native_p),
    "     Other", as.factor(other_race_n), as.factor(other_race_p),
    "Birthplace", row_us_born_n, row_us_born_p,
    "     Born in the US", as.factor(us_born_n), as.factor(us_born_p),
    "     Not Born in the US", as.factor(not_us_born_n), as.factor(not_us_born_p)
    )

# now we can use the built in APA table function to format that table

table_1_caption <- "N = 96. Participants were 18.91 years old on average (SD = 1.15)"

apa_table(dem_table, 
          caption = "Sociodemographic Characteristics of Participants", 
          align = NULL, 
          digits = 2, 
          note = table_1_caption)
```

Participants were recruited through a multicultural orientation event, with the first wave of data collection occurring during their first semester of college and the final wave occurring during the second semester of their sophomore year. Participants who participated in at least three waves of the survey were included in analyses.

## Materials

### College Major Pathway

College major pathways were coded based on participant self-reported narratives about their path to choosing a college major. Students were assigned one of five pathways: Uncertain (uncertain of their major from T1 to T4), Discovery (undecided at T1, chose major by T4), Redirect (changed major between T1 and T4), Solidification (unsure about initial major at T1, certain about same major at T4), and Certain (certain about same major from T1 to T4).

### Locus of Control

Locus of control was measured through the Internal Locus of Control subscale of the 20-item version of the Multi-Measure Agentic Personality Scale (MAPS20) [@côté2016]. Participants responded to five items on a 5-point Likert scale ranging from 1 (Strongly Disagree) to 5 (Strongly Agree), with higher scores indicating higher internal locus of control.

## Procedure

## Data analysis

We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
