---
title             : "Changes in Locus of Control Orientation Across College Major Pathways
"
shorttitle        : "Locus of Control and Major Pathways"

author: 
  - name          : "Varsha Venkatesh"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "75 E River Pkwy, Minneapolis, MN 55455"
    email         : "venka288@umn.edu"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Moin Syed"
    affiliation   : "1"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "University of Minnesota"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library("DescTools")
library("dplyr")
library("summarytools")
r_refs("r-references.bib")

dat <- read.csv("/Users/varshaschool/Documents/loc_pathways/Data/loc_pathways_simulated_data_2025-11-08.csv")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants
Participants were `r nrow(dat)` undergraduate students enrolled at a large, public university in the United States. They ranged in age from `r min(dat$age_sim)` to `r max(dat$age_sim)` years old (*M* = `r mean(dat$age_sim)`, *SD* = `r sd(dat$age_sim)`). The majority of participants identified as `r Mode(dat$race_sim)` (`r (attr(Mode(dat$race_sim), "freq"))/(length(dat$race_sim)) * 100`%) and `r tolower(Mode(dat$gender_sim))` (`r (attr(Mode(dat$gender_sim), "freq"))/(length(dat$gender_sim)) * 100`%). Additionally, the majority (`r (attr(Mode(dat$us_born_sim), "freq"))/(length(dat$us_born_sim)) * 100`%) of participants were born in the United States. A detailed summary of participant demographics is available in Table 1.

```{r descriptive table, echo=FALSE}
#Gender
male_n <- sum(dat$gender_sim == "Male")
male_p <- (male_n/nrow(dat) * 100)
female_n <- sum(dat$gender_sim == "Female")
female_p <- (female_n/nrow(dat) * 100)
nb_n <- sum(dat$gender_sim == "Nonbinary")
nb_p <- (nb_n/nrow(dat) * 100)



dem_table <- tribble(
    ~Characteristic, ~n, ~`%`,
    "Male", male_n, male_p,
    "Female", female_n, female_p,
    "Nonbinary", nb_n, nb_p,
    )

# now we can use the built in APA table function to format that table

apa_table(dem_table,
          caption = "Sociodemographic Characteristics of Participants",
          align = NULL, 
          digits = 2,
          note = "N = 96. Participants were 18.91 years old on average (SD = 1.15)")
```

with the first wave of data collection occurring during their first semester of college and the final wave occurring during the second semester of their sophomore year.

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
